---
title: "Educational Attainment Report"
author: "PPOL 5202"
format: html
editor: visual
params:
  state: "IL"
  county: "COOK"
---

# How does `r params$county` rank in educational attainment?

```{r}
## add code to pull state name and county name from parameters
library(stringr)

state_name = params$state
county_name = params$county
```

Here, we are focusing on the following county within `r params$state`: `r params$county`.

This table shows where `r params$county`` is ranked among `r params$state` in terms of rates of college degrees. We see that the county is ranked \[\] out of \[add R code for total counties\] counties in the state.

```{r}
## add code to create table 
## may want to use kable() from package kableExtra
library(kableExtra)

head(midwest, 20)
```

```{r}
# Get focal county statistics
il_dt = midwest %>%
  filter(state == "IL") 

# Rank the counties based on popdensity
il_dt$rank = rank(-il_dt$percollege)  # Use negative sign to rank in descending order

# Arrange the table by rank
ranked_table = il_dt %>%
  arrange(rank) %>%   # Arrange by rank (highest to lowest)
  select(county, percollege, rank)  # Select relevant columns

# 4. Display the ranked table
print(ranked_table)
```

This graph presents a visual description of where \[add R code for county name\] falls along rates of college degrees, along with two other statistics: rates of high school diplomas and rates of professional school completion.

```{r}
## add code to create plot 


```

## Summary

```{r}
## hint: for this part, you may want to modify the function to return both the plot and the input data for the plot 


```

We see that, for these statistics:

-   \[county name\] has \[higher or lower\] rates of high school diplomas than the state mean
-   \[county name\] has \[higher or lower\] rates of college degrees than the state mean
-   \[county name\] has \[higher or lower\] rates of professional school degrees than the state mean
