---
title: "Educational Attainment Report"
author: "PPOL 5202"
format: html
editor: visual
params:
  state: "IL"
  county: "COOK"
---

# How does \[county name\] rank in educational attainment?

```{r}
## add code to pull state name and county name from parameters
state = params$state
county = params$county
```

Here, we are focusing on the following county within `r state`: `r county`.

This table shows where `r county` is ranked among `r state`\] in terms of rates of college degrees. We see that the county is ranked \[add R code for county rank\] out of \[add R code for total counties\] counties in the state.

```{r}
## add code to create table 
## may want to use kable() from package kableExtra

```

This graph presents a visual description of where \[add R code for county name\] falls along rates of college degrees, along with two other statistics: rates of high school diplomas and rates of professional school completion.

```{r}
## add code to create plot 


```

## Summary

```{r}
## hint: for this part, you may want to modify the function to return both the plot and the input data for the plot 


```

We see that, for these statistics:

-   \[county name\] has \[higher or lower\] rates of high school diplomas than the state mean
-   \[county name\] has \[higher or lower\] rates of college degrees than the state mean
-   \[county name\] has \[higher or lower\] rates of professional school degrees than the state mean
