---
title: "Educational Attainment Report"
author: "PPOL 5202"
format: html
editor: visual
params:
  stateabbr: "IL"
  county: "COOK"
---

# How does `R str_to_title(params$county)` rank in educational attainment?

```{r}
#| echo: false

library(stringr)
library(ggplot2)
library(tidyverse)

full_state <- state.name[state.abb == params$stateabbr]
countyname = str_to_title(params$county)


sortedstatedata = midwest %>% filter(state == params$stateabbr) %>% arrange(desc(percollege)) 


library(tidyverse)
state_name = params$state
county_name = params$county

state_name
county_name

table_data = midwest %>% 
        filter(state == state_name)%>%
        arrange(desc(percollege))

county_ranking = which(table_data$county == county_name)
```

ksl84
Here, we are focusing on the following county within `R full_state`: `R countyname`.

This table shows where `R countyname` is ranked among `R full_state` in terms of rates of college degrees. We see that the county is ranked \[add R code for county rank\] out of \[add R code for total counties\] counties in the state.


```{r}
## add code to create table 
## may want to use kable() from package kableExtra
#install.packages("kableExtra")

library(kableExtra)



kable(table_data) %>%
  kable_styling() %>%
  row_spec(which(table_data$county == county_name), bold = T, color = "white", background = "blue")
```

This graph presents a visual description of where \[add R code for county name\] falls along rates of college degrees, along with two other statistics: rates of high school diplomas and rates of professional school completion.

```{r}
## add code to create plot 


```

## Summary

```{r}
## hint: for this part, you may want to modify the function to return both the plot and the input data for the plot 


```

We see that, for these statistics:

-   \[county name\] has \[higher or lower\] rates of high school diplomas than the state mean
-   \[county name\] has \[higher or lower\] rates of college degrees than the state mean
-   \[county name\] has \[higher or lower\] rates of professional school degrees than the state mean
