---
title: "ggplot2 automation"
format: html
editor: visual
execute: 
  warning: false
  message: false
---

# Automation

In this activity, you'll work to produce the following:

1)  A template that can be used to compare a focal county's demographic statistics to other counties in that state

2)  Use that template to create a separate report for 3 distinct counties

## Exercise 1: create a plot for a single county

-   Use the `midwest` data
-   Focus on `county == "COOK"` within `state == "IL"` as the focal county
-   Create a bar graph where the bars represent the mean value of the following statistics for all the remaining counties in Illinois (so counties minus Cook):
    -   `perchsd`: percent with a high school diploma
    -   `percollege`: percent with a college degree
    -   `percprof`: percent with a professional school degree
-   Adds a dot + label to that bar graph with the focal county's own statistics (at the rough height of where that statistic falls along the `y axis`) compared to the mean value

```{r}
## your code here 

```

## Exercise 2: generalize to create a county comparison function

Write a function that generalizes the plot from exercise 1 by: - Taking in two parameters: county + state - Finding the remaining counties in that state - Creating the same plot as above (making sure to label the county-specific rate appropriately with the county's name) - Test the function on a few different counties from different states

```{r}
## your code here

```

## Exercise 3: use that function to create a county-specific report

-   Modify the `county_template.qmd` file to add in the missing components

-   Write and execute a bash script to generate reports for the following counties within Illinois:

    -   "COOK"
    -   "CHAMPAIGN"
    -   "LAKE"

    See other file for solutions
