---
title: "Geospatial data in Python"
format: html
editor: visual
execute: 
  warning: false
  message: false
---

# Setup

```{r}
library(reticulate)

```

```{python}
import pandas as pd
import numpy as np
import geopandas as gpd
import matplotlib.pyplot as plt
from matplotlib import colormaps
from pygris import states
import folium
import branca.colormap as cm
from branca.colormap import linear

```

# Activity 1: Choropleth map with one poll's data

-   Load the `1976-2020-president.csv` file
-   Filter to:
    -   `year` = 2020
    -   `party_simplified` is REPUBLICAN
-   Create a new column `vote_share` defined as `candidatevotes`/`totalvotes`
-   Use `pygris` to download state polygons (set the year to be 2020)
-   Use `folium` to create a choropleth map with the Republican vote share for 2020 for each state.
    -   Hint: I would recommend merging the data on `STATEFP` prior to plotting. For that, you may need to add leading zeros to the two digit state fips codes. See here for a useful reference: https://pandas.pydata.org/docs/reference/api/pandas.Series.str.pad.html

```{python}

```

```{python}



```

## Activity 2: add a hoverable popup to the plot

For each state, add a hoverable popup with:

- The state name

- The republican vote share (rounded to 2 digits)

```{python}



```

## Activity 3: switch to R and create the same choropleth map using leaflet

-   See here for some example code: https://rstudio.github.io/leaflet/articles/choropleths.html
-   *Hint*: you'll need to reload the presidential results data and use `tigris` to get the state polygons

```{r}
library(tidyverse) 
library(stringr)
library(leaflet)
library(leaflet.extras)
library(tigris)
library(sf)

```
