---
title: "Educational Attainment Report"
author: "PPOL 5202"
format: html
editor: visual
params:
  state: "IL"
  county: "COOK"
execute:
  warning: false
  message: false 
---

```{r}
#| echo: false
#| output: false
library(stringr)
library(tidyverse)

county_name <- paste(str_to_title(params$county), "County")
state_fullname <- state.name[state.abb == params$state]

## find ranking
rank_allcounties <- midwest %>% filter(state == params$state) %>%
          arrange(desc(percollege)) %>%
          mutate(rank = 1:n()) %>%
          select(county, rank, percollege) 

rank_focalcounty <- rank_allcounties %>% filter(county == params$county) %>% pull(rank)

head(rank_allcounties)
rank_focalcounty

```


# How does `r county_name` rank in educational attainment?

Here, we are focusing on the following county within `r state_fullname`: `r county_name`.

This table shows where `r county_name` is ranked among `r state_fullname` in terms of rates of college degrees. We see that the county is ranked `r sprintf("%sth", rank_focalcounty)` out of `r length(unique(rank_allcounties$county))` counties in the state.

```{r}
#| echo: false
#| output: true
library(kableExtra)
rank_allcounties %>%
  kable(booktabs = T) %>%
  kable_styling() %>%
  row_spec(rank_focalcounty, bold = T, color = "white", background = "green")

```

This graph presents a visual description of where `r county_name` falls along rates of college degrees, along with two other statistics: rates of high school diplomas and rates of professional school completion.

```{r}
#| echo: false
#| output: true
## add code to create plot 
library(ggplot2)

comparison_plot <- function(state_name, county_name){
  
  stats_comparison <- rbind.data.frame(midwest %>%
            filter(state == state_name & county != county_name) %>%
            summarise(hsd = mean(perchsd), college =  mean(percollege), 
                      professional = mean(percprof)) %>%
            mutate(category = "other"),
            midwest %>%
            filter(state == state_name & county == county_name) %>%
            summarise(hsd = mean(perchsd), college =  mean(percollege), 
                      professional = mean(percprof)) %>%
            mutate(category = county_name)) %>%
            reshape2::melt(., id.vars = "category") %>%
            mutate(variable_descriptive = 
            factor(case_when(variable == "hsd" ~ "High school degree",
                      variable == "college" ~ "College degree",
                      variable == "professional" ~ "Professional school"),
                levels = c("High school degree",
                           "College degree",
                           "Professional school"),
                ordered = TRUE)) 
  
  stats_toplot <- stats_comparison %>%
          left_join(stats_comparison %>% filter(category == "other") 
                  %>% rename(mean_value = value) %>%
                  select(variable, mean_value),
                  by = "variable") 
  
  p <- ggplot(stats_toplot %>% filter(category == "other"), 
       aes(x = variable_descriptive, 
                            y = value)) +
  geom_bar(stat = "identity", color = "black", 
           fill = "wheat4") +
  xlab("Level of educational attainment") +
  ylab(sprintf("Percentage in county (mean across other\n%s counties)",
      state.name[state.abb == state_name])) +
  geom_point(data = stats_toplot %>% filter(category != "other"),
             aes(x = variable_descriptive, 
                 y = value),
             color = "red", size = 4) +
  geom_label(data = stats_toplot %>% filter(category != "other"),
             aes(x = variable_descriptive, 
                 y = value, 
                 label = sprintf("%s county's rate: %.2f\nMean rate: %.2f",
                                 str_to_title(county_name),
                                 value, 
                                 mean_value)),
          color = "red",
          size = 3,
          nudge_y = 5) +
  ylim(0, 100)
  
  return(list(plot = p, comparison_df = stats_toplot))
  
}

p_and_summary <- comparison_plot(state_name = params$state, county_name = params$county)
p_and_summary$p

```


# Summary 

```{r}
#| echo: false
#| output: false
p_and_summary$comparison_df


compare_tomean <- p_and_summary$comparison_df %>%
                      filter(category != "other") 

hs_compare <- ifelse(compare_tomean %>% filter(variable == "hsd") %>% pull(mean_value) >
                    compare_tomean %>% filter(variable == "hsd") %>% pull(value),
                    "lower", "higher") 
c_compare <- ifelse(compare_tomean %>% filter(variable == "college") %>% pull(mean_value) >
                    compare_tomean %>% filter(variable == "college") %>% pull(value),
                    "lower", "higher") 

prof_compare <- ifelse(compare_tomean %>% filter(variable == "professional") %>% pull(mean_value) >
                    compare_tomean %>% filter(variable == "professional") %>% pull(value),
                    "lower", "higher") 

c_compare
```


We see that, for these statistics:

-   `r county_name` has `r hs_compare` rates of high school diplomas than the state mean
-   `r county_name` has `r c_compare` rates of college degrees than the state mean
-   `r county_name` has `r prof_compare` rates of professional school degrees than the state mean
